\bf Title here

### 2020-07-20 paper/paper.Rmd
new version of plot definition

Overall, the first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function. The second question can be addressed via visualising the rotating high dimensional space. This seems to be a hard task to see beyond 3D, but the fact is that we never really see all the six faces of a cube at one time. The way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically. Thus we can do a similar rotation on the screen to precieve even higher dimension. [there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper]. Thus animated visualisation is needed to preceive the optimisation path in the searching space.

The plot design of visual diagnostics depends on the characteristics of the variables to plot. If the searching points are of interest [point geom is not good; need summarise. While updating points are more manageable -> point geom]. However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point. [stratefy may help? - may need more work here.]

@@

Using the notation we defined earlier, the plot can be defined as plotting

$$x = \{1, 2, \cdots J + 1\}$$

against the five point summary of each $j^*$ on the y-axis:

$$y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}$$

where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$. $\mathds{1}(.)$ is the identity operator.

Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it doesn't report the number of point in each box. We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`. Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry, which is achieved via the `cutoff` argument. A line geometry is also added to link the points with the largest index value in each `tries`. This helps to visualise the improvement made by each `tries`.

@@

The number of interpolation points is much managable than the searching points and always shows a pattern of gradual increase in the index value. Hence a line plot with interpolating order on the x-axis will be desirable. Using notation, the x-axis is a sequence $$x = \{1, 2, \cdots,\sum_{i=1}^Jk_i\}$$ and y-axis are the index values defined by $$y = \{I_{jlk}: k = x\}$$

@@

Polishing points form on itself a global structure with candidate bases and output bases. If we're allowed to abuse the notation a bit, we denote $$A_{J, \ast, \sum_{i = 1}^J k_i + 1}$$ as the first polishing basis and $$A_{J, \ast, \sum_{i = 1}^J k_i + M}$$ as the last polishing basis, where $$M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})$$

Then the index plot incorporating both the interpolating points and polishing points can be defined as

$$x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}$$
$$y = \{I_{jlk}: k = x\}$$

@@

In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are usually of dimension $p \times d$ and hence can't be visualised in a 2D plot. An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA). We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows. This gives a matrix of dimension $n$ by $(p*d)$. Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$. By definition, the principal components maximise the variance of the projection bases in the reduced space. Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$. Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix

$$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$

has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$. The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$

The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object. THis means all the iterators and interesting other variables defined in the global object can be binded with the these two projections and can be mapped to other aesthetics (i.e. color).

In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution. [need to think further about how different type of variables can be shown]. While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation. Geozoo is an option. It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.

>>>>

The first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function. The second question can be addressed via visualising the rotating high dimensional space or its projection on the reduced 2D space. Thus animated visualisation is needed to preceive the optimisation path in the searching space.

<!-- [This seems to be a hard task to see the rotation of high dimensional space, but the way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically. Thus we can do a similar rotation on the screen to precieve even higher dimension. there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].  -->

Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation. To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.

<!-- The plot design of visual diagnostics depends on the characteristics of the variables to plot. If the searching points are of interest [point geom is not good; need summarise. While updating points are more manageable -> point geom]. However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point. [stratefy may help? - may need more work here.] -->

@@

Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it doesn't report the number of point in each box. We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`. Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry, which is achieved via the `cutoff` argument. A line geometry is also added to link the points with the largest index value in each `tries`. This helps to visualise the improvement made by each `tries`. With gramma of graphics [@wickham2010layered], the plot for exploring index value can be defined in three layers as follows:

- Layer 1: boxplot geom
  - data: group by `tries` and filter the observations in the group that have count greater than `cutoff = 15`.
  - x: `tries` is mapped to the x-axis
  - y: the index value after statistical transformation $y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}$ is mapped to the y-axis where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$. $\mathds{1}(.)$ is the identity operator.

- Layer 2: point geom
  - data: group by `tries` and filter the observations in the group that have count less than `cutoff = 15`.
  - x: `tries` is mapped to the x-axis
  - y: `index_val` is mapped to the y-axis

- Layer 3: line geom
  - data: filter the points with the highest index value in each `tries`
  - x: `tries` is mapped to the x-axis
  - y: `index_val` is mapped to the y-axis

@@

Sometimes, it is more interesting to explore the index value on the interpolation path and the plot definition is slightly varied.

- Layer 1: point geom
  - data: filter the observations with `info` being `interpolation` and mutate `id` to be the row number of the subsetted tibble
  - x: `id` is mapped to the x-axis
  - y: `index_val` is mapped to the y-axis

- Layer 2: line geom
  - using line goemetry for the same data and aesthetics

@@

Slight variation of plot definition due to the addition of polishing points is as follows:

- Layer 1: point geom
  - data: filter the observations with `info` being `interpolation`; bind the global object from optimisation and interpolation and form polishing; mutate `id` to be the row number of the binded tibble.
  - x: `id` is mapped to the x-axis
  - y: `index_val` is mapped to the y-axis 
  - color: `method` is mapped to the color aesthetic

- Layer 2: line geom
  - using line goemetry for the same data and aesthetics

<!-- Polishing points form on itself a global structure with candidate bases and output bases. If we're allowed to abuse the notation a bit, we denote $$A_{J, \ast, \sum_{i = 1}^J k_i + 1}$$ as the first polishing basis and $$A_{J, \ast, \sum_{i = 1}^J k_i + M}$$ as the last polishing basis, where $$M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})$$ -->

<!-- Then the index plot incorporating both the interpolating points and polishing points can be defined as  -->

<!-- $$x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}$$ -->
<!-- $$y = \{I_{jlk}: k = x\}$$ -->

@@

In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are usually of dimension $p \times d$ and hence can't be visualised in a 2D plot. An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA). The visualisation can thus be defined as

<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows. This gives a matrix of dimension $n$ by $(p*d)$. Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$. By definition, the principal components maximise the variance of the projection bases in the reduced space. Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$. Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix    -->

<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->

<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$. The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->

<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object. THis means all the iterators and interesting other variables defined in the global object can be binded with the these two projections and can be mapped to other aesthetics (i.e. color). -->

- Layer 1: point geom
  - data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
  - x: the projected basis on the first principal component
  - y: the projected basis on the second principal component
  - color: an variable of interest is mapped onto color

<!-- In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution. [need to think further about how different type of variables can be shown].  -->

While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation. Geozoo is an option. It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.

### 2020-07-20 paper/paper.tex
new version of plot definition

Overall, the first question can be answered using a static plot with
x-axis showing the progression of the optimisation and y-axis showing
the value of the objective function. The second question can be
addressed via visualising the rotating high dimensional space. This
seems to be a hard task to see beyond 3D, but the fact is that we never
really see all the six faces of a cube at one time. The way humans
understand the dimension of a 3D object is either through rotating the
physical 3D object or using shade and line type to annotate the 3D
object in a paper or electronically. Thus we can do a similar rotation
on the screen to precieve even higher dimension. {[}there is likely to
be a learning curve - remember when learning geometry in school, it
takes a while to get used to see 3D cube on a paper{]}. Thus animated
visualisation is needed to preceive the optimisation path in the
searching space.

The plot design of visual diagnostics depends on the characteristics of
the variables to plot. If the searching points are of interest {[}point
geom is not good; need summarise. While updating points are more
manageable -\textgreater{} point geom{]}. However, one must realise that
with hundreds or thousands of searching points, exploring the sample
space in animation could be slow and this is because of the time it
takes to render hundreds point. {[}stratefy may help? - may need more
work here.{]}

@@

Using the notation we defined earlier, the plot can be defined as
plotting

\[x = \{1, 2, \cdots J + 1\}\]

against the five point summary of each \(j^*\) on the y-axis:

\[y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}\]

where \(q\) takes one of \(0, 25, 50, 75, 100\) and \(Q_q(x)\) finds the
qth-quantile for the vector \(x\). \(\mathds{1}(.)\) is the identity
operator.

Boxplot is a suitable candidate that provides five points summary of the
data, however it has one drawback: it doesn't report the number of point
in each box. We may risk losing information on how many points it takes
to find the target basis by displaying the boxplot alone for all
\texttt{tries}. Thus, the number of point for each \texttt{tries} is
displayed at the bottom of each box and we provide options to switch
\texttt{tries} with small number of points to a point geometry, which is
achieved via the \texttt{cutoff} argument. A line geometry is also added
to link the points with the largest index value in each \texttt{tries}.
This helps to visualise the improvement made by each \texttt{tries}.

@@

The number of interpolation points is much managable than the searching
points and always shows a pattern of gradual increase in the index
value. Hence a line plot with interpolating order on the x-axis will be
desirable. Using notation, the x-axis is a sequence
\[x = \{1, 2, \cdots,\sum_{i=1}^Jk_i\}\] and y-axis are the index values
defined by \[y = \{I_{jlk}: k = x\}\]

@@

Polishing points form on itself a global structure with candidate bases
and output bases. If we're allowed to abuse the notation a bit, we
denote \[A_{J, \ast, \sum_{i = 1}^J k_i + 1}\] as the first polishing
basis and \[A_{J, \ast, \sum_{i = 1}^J k_i + M}\] as the last polishing
basis, where
\[M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})\]

Then the index plot incorporating both the interpolating points and
polishing points can be defined as

\[x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}\]
\[y = \{I_{jlk}: k = x\}\]

@@

In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
are usually of dimension \(p \times d\) and hence can't be visualised in
a 2D plot. An option to explore the searching space of these bases is to
explore a reduced space via principal component analysis (PCA). We first
flat each projection bases \(\mathbf{A}\) to a row vector and stack them
by rows. This gives a matrix of dimension \(n\) by \((p*d)\). Centering
and scaling are usually done before performing PCA and we denotes the
pre-processed matrix as \(\tilde{\mathbf{A}}_{n \times (p \times d)}\).
By definition, the principal components maximise the variance of the
projection bases in the reduced space. Using lagrange multipler, this is
equivalent to find the eigenvector of the variance-covariance matrix:
\(\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}\). Decomposing
\(\tilde{\mathbf{A}}\) using singular value decomposition (SVD) gives
\[\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}\] and
hence the variance-covariance matrix

\[\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}\]

has eigenvector
\(\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]\).
The projection mapping to the x- and y-axis uses the first two principal
components and hence
\[\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2\]

The resulting two vectors are of dimension \(n \times 1\), which is
compatible with the dimension of the global object. THis means all the
iterators and interesting other variables defined in the global object
can be binded with the these two projections and can be mapped to other
aesthetics (i.e.~color).

In this way, the first two principle components will take up the x and y
axis and all the other information will be mapped using other aesthetic
attribution. {[}need to think further about how different type of
variables can be shown{]}. While explore the reduced space is an initial
attempt to understand the searching space, there are existing technology
for rotating a higher dimensional space for visualisation. Geozoo is an
option. It generates random points on the high dimensional space and we
can overlay it with the points on the optimisation path to visualise the
spread of it on the high-D sphere.

>>>>

The first question can be answered using a static plot with x-axis
showing the progression of the optimisation and y-axis showing the value
of the objective function. The second question can be addressed via
visualising the rotating high dimensional space or its projection on the
reduced 2D space. Thus animated visualisation is needed to preceive the
optimisation path in the searching space.

Since the global object is already tidy, not further tidying steps is
needed, while certain data wrangling steps
\citep{wickham2016rfordatascience} are still needed to transform the
global object into desirable format for one particular visualisation. To
emphasize on this good practice of data analysis, we will describe the
transformation steps needed for each diagnostic plots before stepping
into visualisation.

@@

Boxplot is a suitable candidate that provides five points summary of the
data, however it has one drawback: it doesn't report the number of point
in each box. We may risk losing information on how many points it takes
to find the target basis by displaying the boxplot alone for all
\texttt{tries}. Thus, the number of point for each \texttt{tries} is
displayed at the bottom of each box and we provide options to switch
\texttt{tries} with small number of points to a point geometry, which is
achieved via the \texttt{cutoff} argument. A line geometry is also added
to link the points with the largest index value in each \texttt{tries}.
This helps to visualise the improvement made by each \texttt{tries}.
With gramma of graphics \citep{wickham2010layered}, the plot for
exploring index value can be defined in three layers as follows:

\begin{itemize}
\tightlist
\item
  Layer 1: boxplot geom

  \begin{itemize}
  \tightlist
  \item
    data: group by \texttt{tries} and filter the observations in the
    group that have count greater than \texttt{cutoff\ =\ 15}.
  \item
    x: \texttt{tries} is mapped to the x-axis
  \item
    y: the index value after statistical transformation
    \(y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}\) is
    mapped to the y-axis where \(q\) takes one of \(0, 25, 50, 75, 100\)
    and \(Q_q(x)\) finds the qth-quantile for the vector \(x\).
    \(\mathds{1}(.)\) is the identity operator.
  \end{itemize}
\item
  Layer 2: point geom

  \begin{itemize}
  \tightlist
  \item
    data: group by \texttt{tries} and filter the observations in the
    group that have count less than \texttt{cutoff\ =\ 15}.
  \item
    x: \texttt{tries} is mapped to the x-axis
  \item
    y: \texttt{index\_val} is mapped to the y-axis
  \end{itemize}
\item
  Layer 3: line geom

  \begin{itemize}
  \tightlist
  \item
    data: filter the points with the highest index value in each
    \texttt{tries}
  \item
    x: \texttt{tries} is mapped to the x-axis
  \item
    y: \texttt{index\_val} is mapped to the y-axis
  \end{itemize}
\end{itemize}

@@

Sometimes, it is more interesting to explore the index value on the
interpolation path and the plot definition is slightly varied.

\begin{itemize}
\tightlist
\item
  Layer 1: point geom

  \begin{itemize}
  \tightlist
  \item
    data: filter the observations with \texttt{info} being
    \texttt{interpolation} and mutate \texttt{id} to be the row number
    of the subsetted tibble
  \item
    x: \texttt{id} is mapped to the x-axis
  \item
    y: \texttt{index\_val} is mapped to the y-axis
  \end{itemize}
\item
  Layer 2: line geom

  \begin{itemize}
  \tightlist
  \item
    using line goemetry for the same data and aesthetics
  \end{itemize}
\end{itemize}

@@

Slight variation of plot definition due to the addition of polishing
points is as follows:

\begin{itemize}
\tightlist
\item
  Layer 1: point geom

  \begin{itemize}
  \tightlist
  \item
    data: filter the observations with \texttt{info} being
    \texttt{interpolation}; bind the global object from optimisation and
    interpolation and form polishing; mutate \texttt{id} to be the row
    number of the binded tibble.
  \item
    x: \texttt{id} is mapped to the x-axis
  \item
    y: \texttt{index\_val} is mapped to the y-axis
  \item
    color: \texttt{method} is mapped to the color aesthetic
  \end{itemize}
\item
  Layer 2: line geom

  \begin{itemize}
  \tightlist
  \item
    using line goemetry for the same data and aesthetics
  \end{itemize}
\end{itemize}

@@

In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
are usually of dimension \(p \times d\) and hence can't be visualised in
a 2D plot. An option to explore the searching space of these bases is to
explore a reduced space via principal component analysis (PCA). The
visualisation can thus be defined as

\begin{itemize}
\tightlist
\item
  Layer 1: point geom

  \begin{itemize}
  \tightlist
  \item
    data: subset the basis of interest and arrange into a matrix format;
    perform PCA on the basis matrix and compute the projected basis on
    the first two principal components; bind the variables from the
    original global object and form a tibble
  \item
    x: the projected basis on the first principal component
  \item
    y: the projected basis on the second principal component
  \item
    color: an variable of interest is mapped onto color
  \end{itemize}
\end{itemize}

While explore the reduced space is an initial attempt to understand the
searching space, there are existing technology for rotating a higher
dimensional space for visualisation. Geozoo is an option. It generates
random points on the high dimensional space and we can overlay it with
the points on the optimisation path to visualise the spread of it on the
high-D sphere.
